Appendix 4: Dynamic species distribution model comparison between INLA, inlabru, and sdmTMB

### 2022-03-17 doc/appendix-binomial-inla.Rmd
Update INLA appendix

There are a couple ways to include the fixed effects with `inlabru`. Unfortunately, using factor variables in the standard equation syntax does not work.
One option is to create a design matrix and add year effects individually to the formula.
The other way to set up the fixed coefficients with inlabru is to use the `factor_full` (or `factor_contrast`) model (seen below within `fac()`).
This option causes the location of the factor fixed-effect parameter estimates to end up in `summary.random$fac`.

@@

```{r sdmTMB-sims}
set.seed(1)
post_sdmTMB <- gather_sims(m, n_sims = 2000)
post_sdmTMB <- post_sdmTMB %>% 
  mutate(variable = 
      factor(`.variable`,
        levels = unique(post_sdmTMB$`.variable`),
        labels = c(seq(2007, 2017, 2), "Range", "AR1 rho", "Random Field SD")
      )
  )
```

@@

```{r inla-densities}
x <- list()
x[[1]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2007) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2007")
x[[2]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2009) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2009")
x[[3]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2011) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2011")
x[[4]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2013) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2013")
x[[5]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2015) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2015")
x[[6]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2017) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2017")
x[[7]] <- inla.smarginal(m_inla$marginals.hyperpar$`Range for i`) %>%
  as_tibble() %>%
  mutate(variable = "range")
x[[8]] <- inla.smarginal(m_inla$marginals.hyperpar$`Stdev for i`) %>%
  as_tibble() %>%
  mutate(variable = "sigma_E")
x[[9]] <- inla.smarginal(m_inla$marginals.hyperpar$`GroupRho for i`) %>%
  as_tibble() %>%
  mutate(variable = "ar1_rho")
post_inla1 <- dplyr::bind_rows(x)
post_inla2 <- post_inla1 %>% mutate(variable = factor(variable,
  levels = unique(post_inla1$variable),
  labels = c(
    "2007", "2009", "2011", "2013", "2015", "2017",
    "Range", "Random Field SD", "AR1 rho"
  )
))
```

>>>>

There are a couple ways to include the factor fixed effects with `inlabru`. Unfortunately, using factor variables in the standard equation syntax does not work.
One option is to create a design matrix and add year effects individually to the formula.
The other way to set up the factor fixed coefficients with inlabru is to use the `factor_full` (or `factor_contrast`) model (seen below within `fac()`).
This option causes the location of the factor fixed-effect parameter estimates to end up in `summary.random$fac`.

@@

```{r sdmTMB-sims}
set.seed(1)
post_sdmTMB <- gather_sims(m, nsim = 2000)
post_sdmTMB <- post_sdmTMB %>% 
  mutate(variable = 
      factor(`.variable`,
        levels = unique(post_sdmTMB$`.variable`),
        labels = c(seq(2007, 2017, 2), "Range", "AR1 rho", "Random Field SD")
      )
  )
```

@@

```{r inla-densities}
x <- list()
x[[1]] <- inla.smarginal(m_inla$marginals.fixed$year_factoryear_factor2007) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2007")
x[[2]] <- inla.smarginal(m_inla$marginals.fixed$year_factoryear_factor2009) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2009")
x[[3]] <- inla.smarginal(m_inla$marginals.fixed$year_factoryear_factor2011) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2011")
x[[4]] <- inla.smarginal(m_inla$marginals.fixed$year_factoryear_factor2013) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2013")
x[[5]] <- inla.smarginal(m_inla$marginals.fixed$year_factoryear_factor2015) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2015")
x[[6]] <- inla.smarginal(m_inla$marginals.fixed$year_factoryear_factor2017) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2017")
x[[7]] <- inla.smarginal(m_inla$marginals.hyperpar$`Range for i`) %>%
  as_tibble() %>%
  mutate(variable = "range")
x[[8]] <- inla.smarginal(m_inla$marginals.hyperpar$`Stdev for i`) %>%
  as_tibble() %>%
  mutate(variable = "sigma_E")
x[[9]] <- inla.smarginal(m_inla$marginals.hyperpar$`GroupRho for i`) %>%
  as_tibble() %>%
  mutate(variable = "ar1_rho")
post_inla1 <- dplyr::bind_rows(x)
post_inla2 <- post_inla1 %>% mutate(variable = factor(variable,
  levels = unique(post_inla1$variable),
  labels = c(
    "2007", "2009", "2011", "2013", "2015", "2017",
    "Range", "Random Field SD", "AR1 rho"
  )
))
```
